### 2017-07-14 natural-deduction.tex
reduction rules

\subsection{Categorical Interpretation of Natural Deductions}

$T$ rules: in the symmetric monoidal closed category of the adjunction model

T\_identity: $id_X:X\rightarrow X$

T\_unitI:

T\_unitE: given $t_1:\Delta\rightarrow Unit$ and $t_2:\Gamma\rightarrow Y$, returns
$\lambda_Y\circ(t_1\otimes t_2):\Gamma\otimes\Delta\rightarrow Unit\otimes Y\rightarrow Y$

T\_tenI: given $t_1:\Gamma\rightarrow X$ and $t_2:\Delta\rightarrow Y$, returns
$t_1\otimes t_2:\Gamma\otimes\Delta\rightarrow X\otimes Y$

T\_tenE: given $t_1:\Gamma\rightarrow X\otimes Y$ and
$t_2:\Delta\otimes X\otimes Y\rightarrow Z$, returns \\
$t_2\circ\e{X\otimes Y,\Delta}\circ t_1\otimes id_\Delta:\Gamma\otimes\Delta\rightarrow(X\otimes Y)\otimes\Delta\rightarrow\Delta\otimes(X\otimes Y)\rightarrow Z$

T\_implI:

T\_implE:

T\_imprI:

T\_imprE:

T\_GI: given $s:FX_1\otimes'...\otimes' FX_n\rightarrow A$, returns \\
$Gs\circ G\m{}^{-1}\circ\eta:X_1\otimes...\otimes X_n\rightarrow GF(X_1\otimes...\otimes X_n)\rightarrow G(FX_1\otimes'...\otimes'FX_n)\rightarrow GA$

S rules: in the Lambek category of the adjunction model

S\_identity: $id_A:A\rightarrow A$

S\_unitI:

S\_unitE:

\subsection{Normaalization and Reduction}