### 2017-08-13 Elle/Elle-output.tex
Elle updated

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{supertabular}
\usepackage{cmll}
\usepackage{txfonts}
\usepackage{color}
\usepackage{tikz}
\usepackage{mathpartir}

\newcommand\encircle[1]{%
  \tikz[baseline=(X.base)] 
    \node (X) [draw, shape=circle, inner sep=-1.5pt] {\strut #1};}

\newcommand{\circop}[1]{\mathop{\encircle{#1}}}

\input{Elle-ott}

\begin{document}

Deriving exchange in Elle comonadicly:

\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny impR}] {
      $$\mprset{flushleft}
    \inferrule* [right={\tiny tenL}] {
      $$\mprset{flushleft}
    \inferrule* [right={\tiny Fl}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny Fl}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny beta}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny tenR}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny ax}] {
                \,
              }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
            }{\texttt{\textcolor{red}{<<no parses (char 10): y0 : Gf B;*** . \mbox{$\mid$}- F y0 : F Gf B >>}}}
            \\
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny ax}] {
                \,
              }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
            }{\texttt{\textcolor{red}{<<no parses (char 10): x0 : Gf A;*** . \mbox{$\mid$}- F x0 : F Gf A >>}}}
          }{\texttt{\textcolor{red}{<<no parses (char 22): y0 : Gf B, x0 : Gf A ;*** . \mbox{$\mid$}- h(F y0) (x) F x0 : h(F Gf B) (x) F Gf A >>}}}
        }{\texttt{\textcolor{red}{<<no parses (char 22): x1 : Gf A, y1 : Gf B ;*** . \mbox{$\mid$}- ex y1 , x1 with y0 , x0 in (h(F y0) (x) F x0) : h(F Gf B) (x) F Gf A >>}}}
      }{\texttt{\textcolor{red}{<<no parses (char 11): x1 : Gf A ;*** y2 : F Gf B \mbox{$\mid$}- let y2 : F Gf B be F y1 in (ex y1 , x1 with y0 , x0 in (h(F y0) (x) F x0)) : h(F Gf B) (x) F Gf A >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** x2 : F Gf A, y2 : F Gf B \mbox{$\mid$}- let x2 : F Gf A be F x1 in (let y2 : F Gf B be F y1 in (ex y1 , x1 with y0 , x0 in (h(F y0) (x) F x0))) : h(F Gf B) (x) F Gf A >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** z : h(F Gf A) (x) F Gf B \mbox{$\mid$}- let z : h(F Gf A) (x) F Gf B be x2 (x) y2 in (let x2 : F Gf A be F x1 in (let y2 : F Gf B be F y1 in (ex y1 , x1 with y0 , x0 in (h(F y0) (x) F x0)))) : (h(F Gf B) (x) F Gf A) >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** . \mbox{$\mid$}- \mbox{$\backslash{}$}l z : h(F Gf A) (x) F Gf B . let z : h(F Gf A) (x) F Gf B be x2 (x) y2 in (let x2 : F Gf A be F x1 in (let y2 : F Gf B be F y1 in (ex y1 , x1 with y0 , x0 in (h(F y0) (x) F x0)))) : (h(F Gf A) (x) F Gf B) -> (h(F Gf B) (x) F Gf A) >>}}}
  \end{math}
\end{center}

Deriving right contraction in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny impR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny tenL}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Fl}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny contrR}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny tenR}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny Fr}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny ax}] {
                    \,
                  }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash  \ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                }{\texttt{\textcolor{red}{<<no parses (char 10): x1 : Gf A;*** . \mbox{$\mid$}- F x1 : F Gf A >>}}}
                \\
                \inferrule* [right={\tiny tenR}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
                  }{\texttt{\textcolor{red}{<<no parses (char 10): y0 : Gf B;*** . \mbox{$\mid$}- F y0 : F Gf B >>}}}
                  \\
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                  }{\texttt{\textcolor{red}{<<no parses (char 10): x0 : Gf A;*** . \mbox{$\mid$}- F x0 : F Gf A >>}}}
                }{\texttt{\textcolor{red}{<<no parses (char 22): y0 : Gf B, x0 : Gf A ;*** . \mbox{$\mid$}- h(F y0) (x) F x0 : h(F Gf B) (x) F Gf A >>}}}
              }{\texttt{\textcolor{red}{<<no parses (char 33): x1 : Gf A, y0 : Gf B, x0 : Gf A ;*** . \mbox{$\mid$}- h(F x1) (x) (h(F y0) (x) F x0) : h(F Gf A) (x) (h(F Gf B) (x) F Gf A) >>}}}
            }{\texttt{\textcolor{red}{<<no parses (char 22): y0 : Gf B, x2 : Gf A ;*** . \mbox{$\mid$}- contrR x2 as x1, x0 in (h(F x1) (x) (h(F y0) (x) F x0)) : h(F Gf A) (x) (h(F Gf B) (x) F Gf A) >>}}}
          }{\texttt{\textcolor{red}{<<no parses (char 11): y0 : Gf B ;*** x3 : F Gf A \mbox{$\mid$}- let x3 : F Gf A be F x2 in (contrR x2 as x1, x0 in (h(F x1) (x) (h(F y0) (x) F x0))) : h(F Gf A) (x) (h(F Gf B) (x) F Gf A) >>}}}
        }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** y1 : F Gf B, x3 : F Gf A \mbox{$\mid$}- let y1 : F Gf B be F y0 in (let x3 : F Gf A be F x2 in (contrR x2 as x1, x0 in (h(F x1) (x) (h(F y0) (x) F x0)))) : h(F Gf A) (x) (h(F Gf B) (x) F Gf A) >>}}}
      }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** z : h(F Gf B) (x) F Gf A \mbox{$\mid$}- let z : h(F Gf B) (x) F Gf A be y1 (x) x3 in (let y1 : F Gf B be F y0 in (let x3 : F Gf A be F x2 in (contrR x2 as x1, x0 in (h(F x1) (x) (h(F y0) (x) F x0))))) : h(F Gf A) (x) (h(F Gf B) (x) F Gf A) >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** . \mbox{$\mid$}- \mbox{$\backslash{}$}l z : h(F Gf B) (x) F Gf A . let z : h(F Gf B) (x) F Gf A be y1 (x) x3 in (let y1 : F Gf B be F y0 in (let x3 : F Gf A be F x2 in (contrR x2 as x1, x0 in (h(F x1) (x) (h(F y0) (x) F x0))))) : (h(F Gf B) (x) F Gf A) -> (h(F Gf A) (x) (h(F Gf B) (x) F Gf A)) >>}}}
  \end{math}
\end{center}

Deriving left contraction in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny impR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny tenL}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Fl}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny contrR}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny tenR}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny tenR}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                  }{\texttt{\textcolor{red}{<<no parses (char 10): x0 : Gf A;*** . \mbox{$\mid$}- F x0 : F Gf A >>}}}
                  \\
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
                  }{\texttt{\textcolor{red}{<<no parses (char 10): y0 : Gf B;*** . \mbox{$\mid$}- F y0 : F Gf B >>}}}
                }{\texttt{\textcolor{red}{<<no parses (char 22): x0 : Gf A, y0 : Gf B ;*** . \mbox{$\mid$}- h(F x0) (x) F y0 : h(F Gf A) (x) F Gf B >>}}}
                \\
                \inferrule* [right={\tiny Fr}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny ax}] {
                    \,
                  }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash  \ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                }{\texttt{\textcolor{red}{<<no parses (char 10): x1 : Gf A;*** . \mbox{$\mid$}- F x1 : F Gf A >>}}}
              }{\texttt{\textcolor{red}{<<no parses (char 33): x0 : Gf A, y0 : Gf B, x1 : Gf A ;*** . \mbox{$\mid$}- (h(F x0) (x) F y0) (x) F x1 : (h(F Gf A) (x) F Gf B) (x) F Gf A >>}}}
            }{\texttt{\textcolor{red}{<<no parses (char 22): x2 : Gf A, y0 : Gf B ;*** . \mbox{$\mid$}- contrL x2 as x0, x1 in ((h(F x0) (x) F y0) (x) F x1) : (h(F Gf A) (x) F Gf B) (x) F Gf A >>}}}
          }{\texttt{\textcolor{red}{<<no parses (char 11): x2 : Gf A ;*** y1 : F Gf B \mbox{$\mid$}- let y1 : F Gf B be F y0 in (contrL x2 as x0, x1 in ((h(F x0) (x) F y0) (x) F x1)) : (h(F Gf A) (x) F Gf B) (x) F Gf A >>}}}
        }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** x3 : F Gf A, y1 : F Gf B \mbox{$\mid$}- let x3 : F Gf A be F x2 in (let y1 : F Gf B be F y0 in (contrL x2 as x0, x1 in ((h(F x0) (x) F y0) (x) F x1))) : (h(F Gf A) (x) F Gf B) (x) F Gf A >>}}}
      }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** z : h(F Gf A) (x) F Gf B \mbox{$\mid$}- let z : h(F Gf A) (x) F Gf B be x3 (x) y1 in (let x3 : F Gf A be F x2 in (let y1 : F Gf B be F y0 in (contrL x2 as x0, x1 in ((h(F x0) (x) F y0) (x) F x1)))) : (h(F Gf A) (x) F Gf B) (x) F Gf A >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** . \mbox{$\mid$}- \mbox{$\backslash{}$}l z : h(F Gf A) (x) F Gf B . let z : h(F Gf A) (x) F Gf B be x3 (x) y1 in (let x3 : F Gf A be F x3 in (let y1 : F Gf B be F y0 in (contrL x2 as x0, x1 in (h(F x0) (x) (h(F y0) (x) F x1))))) : (h(F Gf A) (x) F Gf B) -> ((h(F Gf A) (x) F Gf B) (x) F Gf A) >>}}}
  \end{math}
\end{center}

Deriving weakening in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny impR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny Fl}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny weak}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny unitR}] {
            \,
          }{\texttt{\textcolor{red}{<<no parses (char 10): . \mbox{$\mid$}- triv ***: Unit >>}}}
        }{\texttt{\textcolor{red}{<<no parses (char 11): x0 : Gf A ;*** . \mbox{$\mid$}- weak x0 in triv : Unit >>}}}
      }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** x1 : F Gf A \mbox{$\mid$}- let x1 : F Gf A be F x0 in weak x0 in triv : Unit >>}}}
    }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** . \mbox{$\mid$}- \mbox{$\backslash{}$}l x1 : F Gf A . let x1 : F Gf A be F x0 in weak x0 in triv : (F Gf A) -> (Unit) >>}}}
  \end{math}
\end{center}

$GF$ is a monad:
\begin{itemize}
\item Deriving $\eta$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny impR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Gr}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Gl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fl}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny Gl}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny ax}] {
                  \,
                }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** x0 : F X \mbox{$\mid$}- x0 : F X >>}}}
              }{\texttt{\textcolor{red}{<<no parses (char 13): x1 : Gf F X ;*** . \mbox{$\mid$}- let x1 : Gf F X be Gf x0 in x0 : F X >>}}}
            }{\texttt{\textcolor{red}{<<no parses (char 3): . ;*** x2 : F Gf F X \mbox{$\mid$}- let x2 : F Gf F X be F x1 in (let x1 : Gf F X be Gf x0 in x0) : F X >>}}}
          }{\texttt{\textcolor{red}{<<no parses (char 18): x3 : Gf F Gf F X ;*** . \mbox{$\mid$}- let x3 : Gf F Gf F X be Gf x2 in (let x2 : F Gf F X be F x1 in (let x1 : Gf F X be Gf x0 in x0)) : F X >>}}}
        }{\ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}      \vdash   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}    \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{X}  }
      }{\texttt{\textcolor{red}{<<no parses (char 15):  . \mbox{$\mid$}- \mbox{$\backslash{}$}l x3 : G***f F Gf F X . Gf (let x3 : Gf F Gf F X be Gf x2 in (let x2 : F Gf F X be F x1 in (let x1 : Gf F X be Gf x0 in x0))) : (Gf F Gf F X) -> (Gf F X) >>}}}
    \end{math}
  \end{center}
\item Deriving $\mu$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny impR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Gr}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fr}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny ax}] {
              \,
            }{\ottmv{x}  \ottsym{:}  \ottnt{X}  \vdash  \ottmv{x}  \ottsym{:}  \ottnt{X}}
          }{\texttt{\textcolor{red}{<<no parses (char 7): x : X ;*** . \mbox{$\mid$}- F x : F X >>}}} 
        }{\ottmv{x}  \ottsym{:}  \ottnt{X}  \vdash   \mathsf{G}  \mathsf{F} \ottmv{x}    \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{X}  }
      }{\texttt{\textcolor{red}{<<no parses (char 13): . \mbox{$\mid$}- \mbox{$\backslash{}$}l x : X*** . Gf F x : (X) -> (Gf F X) >>}}}
    \end{math}
  \end{center}
\end{itemize}

The monad $GF$ is strong:
\begin{itemize}
\item Deriving the tensorial strength $\tau$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny impR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny tenL}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny beta}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Gr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny Gl}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny Fl}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny tenR}] {
                      $$\mprset{flushleft}
                      \inferrule* [right={\tiny ax}] {
                        \,
                      }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \vdash  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}}
                      $$\mprset{flushleft}
                      \inferrule* [right={\tiny ax}] {
                        \,
                      }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}  \vdash  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}}
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}  \vdash  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \ottnt{X}   \otimes  \ottnt{Y}}
                  }{\texttt{\textcolor{red}{<<no parses (char 17): x0 : X , y0 : Y ;*** . \mbox{$\mid$}- F (x0 (x) y0) : F (h(X) (x) Y) >>}}}
                }{\texttt{\textcolor{red}{<<no parses (char 8): x0 : X ;*** y1 : F Y \mbox{$\mid$}- let y1 : F Y be F y0 in F (x0 (x) y0) : F (h(X) (x) Y) >>}}}
              }{\texttt{\textcolor{red}{<<no parses (char 22): y2 : Gf F Y , x0 : X ;*** . \mbox{$\mid$}- let y2 : Gf F Y be Gf y1 in (let y1 : F Y be F y0 in F (x0 (x) y0)) : F (h(X) (x) Y) >>}}}
            }{\ottmv{y_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{Y}    \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \vdash   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
          }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{Y}    \vdash   \mathsf{ex}\, \ottmv{y_{{\mathrm{3}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{2}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
        }{\ottmv{z}  \ottsym{:}   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}    \vdash   \mathsf{let}\, \ottmv{z}  :   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{1}}}}  \otimes  \ottmv{y_{{\mathrm{3}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{3}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{2}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
      }{\texttt{\textcolor{red}{<<no parses (char 15): . \mbox{$\mid$}- \mbox{$\backslash{}$}l z : h(X***) (x) Gf F Y . let z : h(X) (x) Gf F Y be x1 (x) y3 in (ex y3 , x1 with y2 , x0 in (Gf (let y2 : Gf F Y be Gf y1 in (let y1 : F Y be F y0 in F (x0 (x) y0))))) : (h(X) (x) Gf F Y) -> (Gf F (h(X) (x) Y)) >>}}}
    \end{math}
  \end{center}
\end{itemize}

\appendix

\section{Full Ott Spec}
\label{sec:full_ott_spec}
\footnotesize
\ottall
% section full_ott_spec (end)

\end{document}