### 2017-08-29 sequent-calculus-ott.tex

The term assignment for sequent calculus of the commutative part of the model, i.e. the SMCC of
the adjunction, is defined in Figure~\ref{fig:elle-smcc}. And the term assignme for the
non-commutative part, i.e. the Lambek category of the adjunction, is defined in
Figure~\ref{fig:elle-lambek}. $\Psi$ and $\Phi$ are contexts for the non-commutative part
and they are lists. $\Gamma$ and $\Delta$ are contexts for the commutative part and they are
multisets, therefore the following exchange rules are implicit.

\begin{mathpar}
\scriptsize
\end{mathpar}

\begin{figure}[!h]
 \scriptsize
  \begin{mdframed}
    \begin{mathpar}
    \end{mathpar}
  \end{mdframed}
\caption{Commutative Part}
\label{fig:elle-nd-smcc}
\end{figure}