### 2017-08-30 natural-deduction-ott.tex
Makefile error fixed

The term assignment for natural deduction of the commutative part of the model, i.e. the SMCC of
the adjunction, is defined in Figure~\ref{fig:elle-nd-smcc}. And the term assignme for the
non-commutative part, i.e. the Lambek category of the adjunction, is defined in
Figure~\ref{fig:elle-nd-lambek}. $\Psi$ and $\Phi$ are contexts for the commutative part
and they are multisets. $\Gamma$ and $\Delta$ are contexts for the mix of the commutative part and
the non-commutative part, and they are lists. Therefore the following exchange rule is implicit.

\begin{figure}[!h]
  \scriptsize
  \begin{mathpar}
    \NDdruleTXXbeta{}
  \end{mathpar}
\end{figure}

\begin{figure}[!h]
  \scriptsize
  \begin{mdframed}
    \begin{mathpar}
      \NDdruleTXXid{} \qquad\qquad \NDdruleTXXunitI{} \qquad\qquad \NDdruleTXXunitE{} \\
      \NDdruleTXXtenI{} \qquad\qquad \NDdruleTXXtenE{} \\
      \NDdruleTXXimpI{} \qquad\qquad \NDdruleTXXimpE{} \qquad\qquad \NDdruleTXXGI{} \\
      \NDdruleSXXbeta{}
    \end{mathpar}
  \end{mdframed}
\caption{Commutative Part}
\label{fig:elle-nd-smcc}
\end{figure}

\begin{figure}[!h]
 \scriptsize
  \begin{mdframed}
    \begin{mathpar}
      \NDdruleSXXid{} \qquad\qquad \NDdruleSXXunitI{} \qquad\qquad \NDdruleSXXunitEOne{} \\
      \NDdruleSXXunitEOne{} \qquad\qquad \NDdruleSXXunitETwo{} \\
      \NDdruleSXXtenI{} \qquad\qquad \NDdruleSXXtenEOne{} \\
      \NDdruleSXXtenETwo{} \qquad\qquad \NDdruleSXXimprI{} \\
      \NDdruleSXXimprE{} \qquad\qquad \NDdruleSXXimplI{} \\
      \NDdruleSXXimplE{} \qquad\qquad \NDdruleSXXGE{} \qquad\qquad \NDdruleSXXFI{} \\
      \NDdruleSXXFE{}
    \end{mathpar}
  \end{mdframed}
\caption{Non-Commutative Part}
\label{fig:elle-nd-lambek}
\end{figure}

We could derive exchange comonadically as follows:

\begin{center}
  \tiny
  \begin{math}
  $$\mprset{flushleft}
  \inferrule* [right={\tiny imprI}] {
    $$\mprset{flushleft}
    \inferrule* [right={\tiny tenE2}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny id}] {
        \,
      }{\NDmv{z}  \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}    \vdash_\mathcal{L}  \NDmv{z}  \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}  }
        $$\mprset{flushleft}
        \inferrule* [right={\tiny FE}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny id}] {
            \,
          }{\NDmv{x_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{A}    \vdash_\mathcal{L}  \NDmv{x_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{A}  }
            $$\mprset{flushleft}
            \inferrule* [right={\tiny FE}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny id}] {
                \,
              }{\NDmv{y_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{B}    \vdash_\mathcal{L}  \NDmv{y_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{B}  }
              \inferrule* [right={\tiny beta}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny FE}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny FI}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny id}] {
                      \,
                    }{\NDmv{y_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{B}   \vdash_\mathcal{C}  \NDmv{y_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{B} }
                  }{\NDmv{y_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{B}   \vdash_\mathcal{L}   \mathsf{F} \NDmv{y_{{\mathrm{0}}}}   \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{B}  }
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny FI}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny id}] {
                      \,
                    }{\NDmv{x_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{A}   \vdash_\mathcal{C}  \NDmv{x_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{A} }
                  }{\NDmv{x_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{A}   \vdash_\mathcal{L}   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{A}  }
                }{\NDmv{y_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{B}   \NDsym{,}  \NDmv{x_{{\mathrm{0}}}}  \NDsym{:}   \mathsf{G} \NDnt{A}   \vdash_\mathcal{L}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}  }
              }{\NDmv{x_{{\mathrm{1}}}}  \NDsym{:}   \mathsf{G} \NDnt{A}   \NDsym{,}  \NDmv{y_{{\mathrm{1}}}}  \NDsym{:}   \mathsf{G} \NDnt{B}   \vdash_\mathcal{L}   \mathsf{ex}\, \NDmv{y_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \NDmv{y_{{\mathrm{0}}}} , \NDmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \NDsym{(}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{)}   \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}  }
            }{\NDmv{x_{{\mathrm{1}}}}  \NDsym{:}   \mathsf{G} \NDnt{A}   \NDsym{,}  \NDmv{y_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{B}    \vdash_\mathcal{L}   \mathsf{let}\,  \mathsf{F} \NDmv{y_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{ex}\, \NDmv{y_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \NDmv{y_{{\mathrm{0}}}} , \NDmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \NDsym{(}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{)}   \NDsym{)}   \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}  }
          }{\NDmv{x_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{A}    \NDsym{,}  \NDmv{y_{{\mathrm{2}}}}  \NDsym{:}   \mathsf{F}  \mathsf{G} \NDnt{B}    \vdash_\mathcal{L}   \mathsf{let}\,  \mathsf{F} \NDmv{x_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{A}   \,\mathsf{be}\, \NDmv{x_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{let}\,  \mathsf{F} \NDmv{y_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{ex}\, \NDmv{y_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \NDmv{y_{{\mathrm{0}}}} , \NDmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \NDsym{(}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{)}   \NDsym{)}   \NDsym{)}   \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}  }
        }{\NDmv{z}  \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \NDmv{z}  :    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{x_{{\mathrm{2}}}}  \triangleright  \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{let}\,  \mathsf{F} \NDmv{x_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{A}   \,\mathsf{be}\, \NDmv{x_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{let}\,  \mathsf{F} \NDmv{y_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{ex}\, \NDmv{y_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \NDmv{y_{{\mathrm{0}}}} , \NDmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \NDsym{(}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{)}   \NDsym{)}   \NDsym{)}   \NDsym{)}   \NDsym{:}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}  }
      }{ \cdot   \vdash_\mathcal{L}   \lambda_r  \NDmv{z}  :    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}   .  \mathsf{let}\, \NDmv{z}  :    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{x_{{\mathrm{2}}}}  \triangleright  \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{let}\,  \mathsf{F} \NDmv{x_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{A}   \,\mathsf{be}\, \NDmv{x_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{let}\,  \mathsf{F} \NDmv{y_{{\mathrm{1}}}}   :   \mathsf{F}  \mathsf{G} \NDnt{B}   \,\mathsf{be}\, \NDmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \NDsym{(}   \mathsf{ex}\, \NDmv{y_{{\mathrm{1}}}} , \NDmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \NDmv{y_{{\mathrm{0}}}} , \NDmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \NDsym{(}    \mathsf{F} \NDmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \NDmv{x_{{\mathrm{0}}}}   \NDsym{)}   \NDsym{)}   \NDsym{)}   \NDsym{)}    \NDsym{:}  \NDsym{(}    \mathsf{F}  \mathsf{G} \NDnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{B}    \NDsym{)}  \rightharpoonup  \NDsym{(}    \mathsf{F}  \mathsf{G} \NDnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \NDnt{A}    \NDsym{)}}
  \end{math}
\end{center}

We also have the three cut rules derivable in the natural deduction:

\begin{figure}[!h]
  \scriptsize
  \begin{mathpar}
    \NDdruleTXXcut{} \qquad\qquad \NDdruleSXXcutOne{} \qquad\qquad \NDdruleSXXcutTwo{}
  \end{mathpar}
\end{figure}

\begin{center}
  \tiny
  \begin{math}
  $$\mprset{flushleft}
  \inferrule* [right={\tiny FI}] {
    {\Phi  \vdash_\mathcal{C}  \NDnt{t}  \NDsym{:}  \NDnt{X}}
  }{\Phi  \vdash_\mathcal{L}   \mathsf{F} \NDnt{t}   \NDsym{:}   \mathsf{F} \NDnt{X} }
  \end{math}
\end{center}