### 2017-11-06 Elle/Elle-output.tex

\documentclass[11pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{supertabular}
\usepackage{cmll}
\usepackage{txfonts}
\usepackage{color}
\usepackage{tikz}
\usepackage{mathpartir}

\newcommand\encircle[1]{%
  \tikz[baseline=(X.base)] 
    \node (X) [draw, shape=circle, inner sep=-1.5pt] {\strut #1};}

\newcommand{\circop}[1]{\mathop{\encircle{#1}}}

\input{Elle-ott}

\begin{document}

Deriving exchange in Elle comonadicly:

\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny imprR}] {
      $$\mprset{flushleft}
    \inferrule* [right={\tiny tenL}] {
      $$\mprset{flushleft}
    \inferrule* [right={\tiny Fl}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny Fl}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny beta}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny tenR}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny ax}] {
                \,
              }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{C}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
            }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}  }
            \\
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny ax}] {
                \,
              }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
            }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}  }
          }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
        }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}   \mathsf{ex}\, \ottmv{y_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
      }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
    }{\ottmv{x_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{,}  \ottmv{y_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
    }{\ottmv{z}  \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{2}}}}  \triangleright  \ottmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
    }{ \cdot   \vdash_\mathcal{L}   \lambda_r  \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   .  \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{2}}}}  \triangleright  \ottmv{y_{{\mathrm{2}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}    \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \rightharpoonup  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
  \end{math}
\end{center}

Deriving right contraction in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny imprR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny tenL}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Fl}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny contrR}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny tenR}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny Fr}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny ax}] {
                    \,
                  }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}  }
                \\
                \inferrule* [right={\tiny tenR}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{C}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
                  }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}  }
                  \\
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                  }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}  }
                }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
              }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
            }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{contrR}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
          }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrR}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{)}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
        }{\ottmv{y_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{,}  \ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrR}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
      }{\ottmv{z}  \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\, \ottmv{y_{{\mathrm{1}}}}  \triangleright  \ottmv{x_{{\mathrm{3}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrR}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
    }{ \cdot   \vdash_\mathcal{L}   \lambda_r  \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}   .  \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\, \ottmv{y_{{\mathrm{1}}}}  \triangleright  \ottmv{x_{{\mathrm{3}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrR}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{1}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{1}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}  \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}    \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}  \rightharpoonup  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{B}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}  \ottsym{)}}
  \end{math}
\end{center}

Deriving left contraction in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny imprR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny tenL}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Fl}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny contrR}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny tenR}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny tenR}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                  }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}  }
                  \\
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny ax}] {
                      \,
                    }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{C}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B} }
                  }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}  }
                }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}  }
                \\
                \inferrule* [right={\tiny Fr}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny ax}] {
                    \,
                  }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A} }
                }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}  }
              }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \ottsym{,}  \ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}  \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{)}  \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
            }{\ottmv{x_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{B}   \vdash_\mathcal{L}   \mathsf{contrL}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}  \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{)}  \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{)}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
          }{\ottmv{x_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \ottsym{,}  \ottmv{y_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrL}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}  \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{)}  \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{)}   \ottsym{)}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
        }{\ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{,}  \ottmv{y_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrL}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}  \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{)}  \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
      }{\ottmv{z}  \ottsym{:}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{3}}}}  \triangleright  \ottmv{y_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrL}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}  \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{y_{{\mathrm{0}}}}   \ottsym{)}  \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}  }
    }{ \cdot   \vdash_\mathcal{L}   \lambda_r  \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   .  \mathsf{let}\, \ottmv{z}  :    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{3}}}}  \triangleright  \ottmv{y_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{3}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{B}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{contrL}\, \ottmv{x_{{\mathrm{2}}}} \,\mathsf{as}\, \ottmv{x_{{\mathrm{0}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{in}\, \ottsym{(}    \mathsf{F} \ottmv{x_{{\mathrm{0}}}}    \triangleright  \ottsym{(}    \mathsf{F} \ottmv{y_{{\mathrm{0}}}}    \triangleright   \mathsf{F} \ottmv{x_{{\mathrm{1}}}}   \ottsym{)}  \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}    \ottsym{:}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \rightharpoonup  \ottsym{(}  \ottsym{(}    \mathsf{F}  \mathsf{G} \ottnt{A}     \triangleright   \mathsf{F}  \mathsf{G} \ottnt{B}    \ottsym{)}  \triangleright   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}}
  \end{math}
\end{center}

Deriving weakening in Elle comonadicly:
\begin{center}
  \tiny
  \begin{math}
    $$\mprset{flushleft}
    \inferrule* [right={\tiny imprR}] {
      $$\mprset{flushleft}
      \inferrule* [right={\tiny Fl}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny weak}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny unitR}] {
            \,
          }{ \cdot   \vdash_\mathcal{L}   \mathsf{trivS}   \ottsym{:}   \mathsf{UnitS} }
        }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{G} \ottnt{A}   \vdash_\mathcal{L}   \mathsf{weak}\, \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\,  \mathsf{trivS}    \ottsym{:}   \mathsf{UnitS} }
      }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G} \ottnt{A}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{weak}\, \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\,  \mathsf{trivS}     \ottsym{:}   \mathsf{UnitS} }
    }{ \cdot   \vdash_\mathcal{L}   \lambda_r  \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   .  \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{F}  \mathsf{G} \ottnt{A}   \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{weak}\, \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\,  \mathsf{trivS}      \ottsym{:}  \ottsym{(}   \mathsf{F}  \mathsf{G} \ottnt{A}    \ottsym{)}  \rightharpoonup  \ottsym{(}   \mathsf{UnitS}   \ottsym{)}}
  \end{math}
\end{center}

$GF$ is a monad:
\begin{itemize}
\item Deriving $\eta$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny imprR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Gr}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Gl}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Fl}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny Gl}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny ax}] {
                  \,
                }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{F} \ottnt{X}   \vdash_\mathcal{L}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}   \mathsf{F} \ottnt{X} }
              }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{X}    \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{:}   \mathsf{F} \ottnt{X} }
            }{\ottmv{x_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}    \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{:}   \mathsf{F} \ottnt{X} }
          }{\ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}      \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}    \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{F} \ottnt{X} }
        }{\ottmv{x_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}      \vdash_\mathcal{C}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}    \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{X}  }
      }{ \cdot   \vdash_\mathcal{C}   \lambda  \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     .  \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{3}}}}  :   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}     \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{2}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{2}}}}  :   \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}    \,\mathsf{be}\,  \mathsf{F}\, \ottmv{x_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{x_{{\mathrm{1}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{X}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{x_{{\mathrm{0}}}}  \,\mathsf{in}\, \ottmv{x_{{\mathrm{0}}}}   \ottsym{)}   \ottsym{)}   \ottsym{)}    \ottsym{:}  \ottsym{(}   \mathsf{G}  \mathsf{F}  \mathsf{G}  \mathsf{F} \ottnt{X}      \ottsym{)}  \multimap  \ottsym{(}   \mathsf{G}  \mathsf{F} \ottnt{X}    \ottsym{)}}
    \end{math}
  \end{center}
\item Deriving $\mu$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny imprR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny Gr}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny Fr}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny ax}] {
              \,
            }{\ottmv{x}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{C}  \ottmv{x}  \ottsym{:}  \ottnt{X}}
          }{\ottmv{x}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{L}   \mathsf{F} \ottmv{x}   \ottsym{:}   \mathsf{F} \ottnt{X} } 
        }{\ottmv{x}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{C}   \mathsf{G}  \mathsf{F} \ottmv{x}    \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{X}  }
      }{ \cdot   \vdash_\mathcal{C}   \lambda  \ottmv{x}  :  \ottnt{X} .  \mathsf{G}  \mathsf{F} \ottmv{x}     \ottsym{:}  \ottsym{(}  \ottnt{X}  \ottsym{)}  \multimap  \ottsym{(}   \mathsf{G}  \mathsf{F} \ottnt{X}    \ottsym{)}}
    \end{math}
  \end{center}
\end{itemize}

The monad $GF$ is strong:
\begin{itemize}
\item Deriving the tensorial strength $\tau$:
  \begin{center}
    \tiny
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\tiny imprR}] {
        $$\mprset{flushleft}
        \inferrule* [right={\tiny tenL}] {
          $$\mprset{flushleft}
          \inferrule* [right={\tiny beta}] {
            $$\mprset{flushleft}
            \inferrule* [right={\tiny Gr}] {
              $$\mprset{flushleft}
              \inferrule* [right={\tiny Gl}] {
                $$\mprset{flushleft}
                \inferrule* [right={\tiny Fl}] {
                  $$\mprset{flushleft}
                  \inferrule* [right={\tiny Fr}] {
                    $$\mprset{flushleft}
                    \inferrule* [right={\tiny tenR}] {
                      $$\mprset{flushleft}
                      \inferrule* [right={\tiny ax}] {
                        \,
                      }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}}
                      $$\mprset{flushleft}
                      \inferrule* [right={\tiny ax}] {
                        \,
                      }{\ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}  \vdash_\mathcal{C}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}}
                    }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}  \vdash_\mathcal{C}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}   \ottnt{X}   \otimes  \ottnt{Y}}
                  }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{Y}  \vdash_\mathcal{L}   \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}   \ottsym{:}   \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)} }
                }{\ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{1}}}}  \ottsym{:}   \mathsf{F} \ottnt{Y}   \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{:}   \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)} }
              }{\ottmv{y_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{Y}    \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{L}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{:}   \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)} }
            }{\ottmv{y_{{\mathrm{2}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{Y}    \ottsym{,}  \ottmv{x_{{\mathrm{0}}}}  \ottsym{:}  \ottnt{X}  \vdash_\mathcal{C}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
          }{\ottmv{x_{{\mathrm{1}}}}  \ottsym{:}  \ottnt{X}  \ottsym{,}  \ottmv{y_{{\mathrm{3}}}}  \ottsym{:}   \mathsf{G}  \mathsf{F} \ottnt{Y}    \vdash_\mathcal{C}   \mathsf{ex}\, \ottmv{y_{{\mathrm{3}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{2}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
        }{\ottmv{z}  \ottsym{:}   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}    \vdash_\mathcal{C}   \mathsf{let}\, \ottmv{z}  :   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{1}}}}  \otimes  \ottmv{y_{{\mathrm{3}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{3}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{2}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{:}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}  }
      }{ \cdot   \vdash_\mathcal{C}   \lambda  \ottmv{z}  :   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}   .  \mathsf{let}\, \ottmv{z}  :   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\, \ottmv{x_{{\mathrm{1}}}}  \otimes  \ottmv{y_{{\mathrm{3}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{ex}\, \ottmv{y_{{\mathrm{3}}}} , \ottmv{x_{{\mathrm{1}}}} \,\mathsf{with}\, \ottmv{y_{{\mathrm{2}}}} , \ottmv{x_{{\mathrm{0}}}} \,\mathsf{in}\, \ottsym{(}   \mathsf{G} \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{2}}}}  :   \mathsf{G}  \mathsf{F} \ottnt{Y}   \,\mathsf{be}\,  \mathsf{G}\, \ottmv{y_{{\mathrm{1}}}}  \,\mathsf{in}\, \ottsym{(}   \mathsf{let}\, \ottmv{y_{{\mathrm{1}}}}  :   \mathsf{F} \ottnt{Y}  \,\mathsf{be}\,  \mathsf{F}\, \ottmv{y_{{\mathrm{0}}}}  \,\mathsf{in}\,  \mathsf{F} \ottsym{(}  \ottmv{x_{{\mathrm{0}}}}  \otimes  \ottmv{y_{{\mathrm{0}}}}  \ottsym{)}    \ottsym{)}   \ottsym{)}   \ottsym{)}   \ottsym{)}    \ottsym{:}  \ottsym{(}   \ottnt{X}   \otimes   \mathsf{G}  \mathsf{F} \ottnt{Y}    \ottsym{)}  \multimap  \ottsym{(}   \mathsf{G}  \mathsf{F} \ottsym{(}   \ottnt{X}   \otimes  \ottnt{Y}  \ottsym{)}    \ottsym{)}}
    \end{math}
  \end{center}
\end{itemize}

\appendix

\section{Full Ott Spec}
\label{sec:full_ott_spec}
\footnotesize
\ottall
% section full_ott_spec (end)

\end{document}