### 2017-11-20 linearity-2018/adjoint-models.tex
Finished the introduction, and moving on to:
The first section on adjoint models, and then the systems.

We first introduce adjoint models, and the two particular adjoint
models that will be used to model the Lambek Calculus with and without
associativity.  However, before we can do this we have to define a
very simple categorical structure known as a magmoidal category with a
unit.
%% \begin{definition}
%%   \label{def:magmoidal-categories}
%%   1
%% \end{definition}