### 2018-09-13 post-linearity-2018/reviews-and-responses.txt
Uncompilable Post Proceeding for Linearity

# Linearity Paper Final Version Revision

May 22, 2018

## -------------------- Review 1 --------------------
PAPER: 3
TITLE: On the Lambek Calculus with an Exchange Modality AUTHORS: Jiaming Jiang, Harley Eades Iii and Valeria de Paiva
Overall evaluation: 1 (weak accept)
----------- Overall evaluation -----------

This paper introduces a type system for (non-symmetric) monoidal
adjunctions between a symmetric monoidal closed category and a
monoidal bi-closed category called CNC logic. It can be considered as
a variant of Benton's LNL logic for symmetric monoidal adjunctions
between a cartesian closed category and a symmetric monoidal closed
category.  This paper is in general clearly written and should be easy
to follow for readers familiar with the categorical models of linear
logic. I did not check the detail of CNC logic (largely because there
is no information on its equational or rewriting theory). However,
since it is based on well-established categorical structures, I
believe that there should be no substantial errors.  My knowledge on
the Lambek Calculus is rather limited, but I think that having a good
categorical semantics and type system for Lambek Calculus should be
useful and makes a good case of technology-transfer from Linear Logic
to the Lambek Calculus. Thus I recommend acceptance.

- Lemma 7:

I think that this result was already in Kock's 1970 original paper.

- Corollary 11:

I think that the statement of this corollary is wrong. As Benton (and
many other people) points out, the category of coalgebras of a linear
exponential comonad on a symmetric monoidal closed category does not
have to be cartesian closed (and we should take some subcategory
instead). The same applies to the setting of this paper.

- Theorem 18:

I am not sure what "a complete moel" means - completeness
w.r.t. derivablity or w.r.t. equational theory? (It seems that the
same question applies to [8].)

- Section 4:

I agree that the Dialectica categories developped here are
interesting, but also wonder if some much easier models could also do
the good job without much technical complication.  For instance,
starting with a strong monoidal functor F:C->D from a symmetric
monoidal category C to a monoidal category D, the monoidal
cocompletion (i.e. taking the presheaf categories) immediately gives
an LAM [C^op,Set] -> [D^op.Set], thus a sound model of CNC logic.

- A work which might be related to this paper is

Masahito Hasegawa:

Linear Exponential Comonads without Symmetry. LINEARITY 2016: 54-63
where the author axiomatizes comonads corresponding to monoidal
adjunctions between a cartesian category and a monoidal category. It
might be interesting to do the same for LAM (which should be something
like "linear exchange comonads").

- On the direction of arrows:

It seems that the direction of arrow types is not in harmony with the
standard Lambek Calculus nor with the usual convention for internal
homs in monoidal bi-closed categories. Is there any reason for this
choice?

### Response to Review 1:
* Will add the reference by Hasegawa.

* Lemma 7

Only found half of it is in Kock's 1970 paper. The full if and only if
version is in Kock's 1972 paper.  As said at the beginning of the 1972
paper: "In [4] (i.e., *Monads on Symmetric Monoidal Closed
Categories*) we proved that a commutative monad on a symmetric
monoidal closed category carries the structure of a symmetric monoidal
monad ([4], Theorem 3.2). We here prove the converse, so that, taken
together, we have: there is a 1-1 correspondence between commutative
monads and symmetric monoidal monads (Theorem 2.3 below)."

**Have added the reference and changed the citation for Lemma 7.**
* Corollary 11

The reviewer thought we were trying to prove that it is cartesian
closed.  The reviewer writes "As Benton (and many other people) points
out, the category of coalgebras of a linear exponential comonad on a
symmetric monoidal closed category does not have to be cartesian
closed (and we should take some subcategory instead)."

HDE: They are actually correct, and it is about closure rather than
carteisan.  I reworded the corollary to be more precise.

**Didn't change anything**
* Theorem 18
The reviewer is not sure what "complete" means.
**Added "(w.r.t. derivability and quational theory)"**
* Direction of Arrows
Not sure why the reviewer thinks the direction of arrows is
wrong. Wikipedia on closed monoidal category uses $\Rightarrow$ for
right adjoint and $\Leftarrow$ for left adjoint. The original Lambek
paper uses \ and /. The page on closed monoidal cateogry in nLab
(https://ncatlab.org/nlab/show/closed+monoidal+category) uses square
brackets, i.e., $[b,c]$ for $b\Rightarrow c$.
**Didn't change anything**

## ----------------------- Review 2 ---------------------
PAPER: 3
TITLE: On the Lambek Calculus with an Exchange Modality AUTHORS: Jiaming Jiang, Harley Eades Iii and Valeria de Paiva
Overall evaluation: 2 (accept)
----------- Overall evaluation -----------

The paper seems as a good contribution to the development of the
research in Lambek Calculus and its extensions and in Non Commutative
Linear Logic. The results seem accurate and well presented. The
authors have a very good experience in this field. It would be
interesting to compare this work with Non Commutative Logic introduced
by Abrusci and Ruet.

### Response to Review 2
* Reference by Abrusci and Ruet:
**Will add the reference.**