### 2018-09-13 post-linearity-2018/sequent.tex

Section for sequent calculus

\begin{mathpar}
\end{mathpar}