### 2018-09-21 post-linearity-2018/sequent.tex
Fixed compilation.  You needed to include:
\input{ElleType-inc}

in main.tex

Section for sequent calculus
\begin{mathpar}
  \SCdruleTXXax{} \and
  \SCdruleTXXunitL{} \and
  \SCdruleTXXunitR{} \and
  \SCdruleTXXbeta{} \and
  \SCdruleTXXcontrR{} \and
  \SCdruleTXXcontrL{} \and
  \SCdruleTXXweak{} \and
  \SCdruleTXXcut{} \and
  \SCdruleTXXcutn{} \and
  \SCdruleTXXtenL{} \and
  \SCdruleTXXtenR{} \and
  \SCdruleTXXimpL{} \and
  \SCdruleTXXimpR{} \and
  \SCdruleTXXGr{}  
\end{mathpar}